---
title: "covid-vignette-api"
author: "Min-Jung Jung"
date: "9/20/2021"
output:   
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reqired pakages

```{r, echo=TRUE, warning = FALSE, message=FALSE, include=FALSE}
library(ggplot2)
library(tidyverse)
library(jsonlite)
library(knitr)
library(httr)
library(tidytext)
```

# JSON Data

JSON, also known as [*Java Script Object Notation*](https://www.json.org/json-en.html) is a text based format for storing and/or transporting data. JSON is used across the internet and databases because of its text based format for storing data. In addition, JSON can represent 2D data, [hierarchical data](https://en.wikipedia.org/wiki/Hierarchical_database_model), and use key-value pairs.

Since JSON is a text based format, we need to load in R packages that can handle the text based format.

# Packages used for reading JSON data in R

Here are R packages that can be used to work with JSON data:

* `rjson`
* `RJSONIO`
* `jsonlite`

For this project, `rjson` and `jsonlite` is used.

Couple resources for understanding the functionality of `rjson` and `jsonlite`:

1. [https://cran.r-project.org/web/packages/rjson/rjson.pdf](https://cran.r-project.org/web/packages/rjson/rjson.pdf)

2. [https://cran.r-project.org/web/packages/jsonlite/jsonlite.pdf](https://cran.r-project.org/web/packages/jsonlite/jsonlite.pdf)

# Contact the Covid Data API

To access the [Covid Data](https://documenter.getpostman.com/view/10808728/SzS8rjbc), we need to get a URL with the name of the table and attributes we want to pull from it.

In addition, I wrote 6 more functions that take different endpoints.


```{r}
base_url = "https://api.covid19api.com"
```

## Function for United States Confirmed Cases from March 1st, 2020 to September 21st, 2021
```{r}
country_Name <- function(){
  country <- GET("https://api.covid19api.com/countries")
  countrylist <- fromJSON(rawToChar(country$content))
  countrylist1 <- as_tibble(data.frame(Country = countrylist$Country, Slug = countrylist$Slug))
  return(countrylist1)
}
```
countryName <- countryName()



```{r}
get_confirmed_cases <- function(country){
     if(country %in% countryName$Slug){
  full_url = paste0(base_url,"/total/country/",country,"/status/confirmed?from=2020-03-01T00:00:00Z&to=2021-09-21T00:00:00Z")
  confirmed_cases_text = content(GET(url=full_url),"text")
  confirmed_cases_json = fromJSON(confirmed_cases_text)
  return(confirmed_cases_json)
     }
        else {
      message <- paste("ERROR: Argument for country was not found in the Slug.", 
                       "Look up countryName to find the country you are looking for and use Slug.")
      stop(message)
    }
}
```



confirmed <- get_confirmed_cases("canada")

confirmed



## Function for United States Death Cases from March 1st, 2020 to September 21st, 2021


```{r}
get_deaths_cases <- function(country){
    if(country %in% countryName$Slug){
  full_url = paste0(base_url,"/total/country/",country,"/status/deaths?from=2020-03-01T00:00:00Z&to=2021-09-21T00:00:00Z")
  deaths_cases_text = content(GET(url=full_url),"text")
  deaths_cases_json = fromJSON(deaths_cases_text)
  return(deaths_cases_json)
    }
        else {
      message <- paste("ERROR: Argument for country was not found in the Slug.", 
                       "Look up countryName to find the country you are looking for and use Slug.")
      stop(message)
    }
}
```


deaths <- get_deaths_cases("india")

deaths





get_deaths_cases("canada")


## Function for United States Recovered Cases from March 1st, 2020 to September 21st, 2021


```{r}
get_recovered_cases <- function(country){
  if(country %in% countryName$Slug){
  full_url = paste0(base_url,"/total/country/", country,"/status/recovered?from=2020-03-01T00:00:00Z&to=2021-09-21T00:00:00Z")
  recovered_cases_text = content(GET(url=full_url),"text")
  recovered_cases_json = fromJSON(recovered_cases_text)
  return(recovered_cases_json)
  }
      else {
      message <- paste("ERROR: Argument for country was not found in the Slug.", 
                       "Look up countryName to find the country you are looking for and use Slug.")
      stop(message)
    }
}
```


recovered <- get_recovered_cases("united-states")
recovered



```{r}
get_cases_bystate <- function(state_name){
  state_name <- tolower(state_name)
  two_word_states = list("new hampshire", "new jersey", "new mexico","new york","north carolina","north dakota","south carolina","south dakota", "distrct of columbia", "puerto rico","Northern Mariana Islands", "Virgin Islands", "Rhode Island")
  if (state_name %in% two_word_states){
     full_url = paste0(base_url,"/dayone/country/united-states/status/confirmed/live?province=",state_name)
     URLencode(full_url)
    covid_cases_by_states_text = content(GET(url=URLencode(full_url)),"text")
    covid_cases_by_states_json = fromJSON(covid_cases_by_states_text)
  }
    else{
    full_url = paste0(base_url,"/dayone/country/united-states/status/confirmed/live?province=",state_name)
    covid_cases_by_states_text = content(GET(url=full_url),"text")
    covid_cases_by_states_json = fromJSON(covid_cases_by_states_text)
  }
  return(covid_cases_by_states_json)
}

```
stateCases <- get_cases_bystate("North Carolina")
stateCases


```{r}
covid_summary_cases <- function(){
   full_url = paste0(base_url,"/summary")
   covid_summary <- GET(url=full_url)
   covid_cases_summary <- fromJSON(rawToChar(covid_summary$content))
   covid_cases_summary1 <- covid_cases_summary$Countries
    return(covid_cases_summary1)
}

```

covidSummary <- covid_summary_cases()
covidSummary


```{r}
get_cases_summary <- function(){
 full_url = paste0(base_url,"/summary")
    covid_cases_by_all_text = content(GET(url=full_url),"text")
    covid_cases_by_all_json = fromJSON(covid_cases_by_all_text)
  return(covid_cases_by_all_json)
}
```
caseSum <- get_cases_summary() 
caseSum



```{r}
get_cases_world <- function(){
 full_url = paste0(base_url,"/world/total")
    covid_cases_world_text = content(GET(url=full_url),"text")
    covid_cases_world_json = fromJSON(covid_cases_world_text)
  return(covid_cases_world_json)
}
```
caseWorld <- get_cases_world()
caseWorld


```{r}
get_live_cases <- function(country){
       if(country %in% countryName$Slug){
   full_url = paste0(base_url,"/live/country/",country,"/status/confirmed/date/2021-08-31T00:00:00Z")
    covid_cases_live_text = content(GET(url=full_url),"text")
    covid_cases_live_json = fromJSON(covid_cases_live_text)
  return(covid_cases_live_json)
       }
      else {
      message <- paste("ERROR: Argument for country was not found in the Slug.", 
                       "Look up countryName to find the country you are looking for and use Slug.")
      stop(message)
    }
}
``` 
liveCase <- get_live_cases("united-states")
liveCase


URLencode()
//Replace space with %20 for url to understand
$new = str_replace(' ', '%20', $your_string);

```{r}
countryName <- country_Name()

confirmed <- get_confirmed_cases("canada")

deaths <- get_deaths_cases("india")

recovered <- get_recovered_cases("united-states")

stateCases <- get_cases_bystate("North Carolina")

caseSum <- get_cases_summary() 

covidSummary <- covid_summary_cases()

caseWorld <- get_cases_world()

country_liveCase <- get_live_cases("united-states")
 

```
# Exploratory Data Analysis

## Contigency Tables/ Numerical Summaries

```{r}
kable(stateCases)

# kable(confirmed)
# kable(deaths)
# kable(recovered)
# kable(caseSum)
# kable(liveCase)
```

## Graphical Summaries


# Bar plot
top_state_09_30_2021 <- country_liveCase %>% filter(Date == "2021-09-30T00:00:00Z") %>% arrange(desc(Active))
top_10_state_09_30_2021 <- top_state_09_30_2021[1:10,]

ggplot(data=top_10_state_09_30_2021, aes(x=Province, y=Active)) +
  geom_bar(stat="identity", fill="green")




  geom_text(aes(label=Cases), vjust=1.6, color="black", size=3.5)+   theme_minimal()+ ggtitle("Top 10 North Carolina Cities Confirmed Covid Cases on September 23rd, 2021.")
  
```{r}
# Bar plot
top_city_09_30_2021 <- stateCases %>% filter(Date == "2021-09-30T00:00:00Z") %>% arrange(desc(Cases))
top_10_city_09_30_2021 <- top_city_09_30_2021[1:10,]

ggplot(data=top_10_city_09_30_2021, aes(x=City, y=Cases)) +
  geom_bar(stat="identity", fill="steelblue")


```

  geom_text(aes(label=Cases), vjust=1.6, color="white", size=3.5)+
stat="identity", theme_minimal()

  + ggtitle("Top 10 North Carolina Cities Confirmed Covid Cases on September 30, 2020.")


# Histogram

top_country_09_30_2021 <- covidSummary %>% filter(Date == "2021-10-01T18:17:16.341Z") %>% arrange(desc(TotalConfirmed))
top_10_country_09_30_2021 <- top_country_09_30_2021[1:20,]

top_10_country_09_30_2021

ggplot(top_10_country_09_30_2021, aes(x=TotalConfirmed)) + geom_histogram(color="darkblue",fill="red",binwidth=1000)+geom_vline(aes(xintercept=mean(TotalConfirmed)),color="blue", linetype="dashed", size=1)+ggtitle("Histogram of Confirmed Covid Cases in top 20 countries on September 30, 2020")

#Blue line is the mean line.


  
```{r}
# Histogram
ggplot(top_city_09_30_2021, aes(x=Cases)) + geom_histogram(color="darkblue",fill="red",binwidth=1000)+geom_vline(aes(xintercept=mean(Cases)),color="blue", linetype="dashed", size=1)+ggtitle("Histogram of Confirmed Covid Cases in North Carolina Cities on November 22nd, 2020")

#Blue line is the mean line.
```