---
title: "covid-vignette-api"
author: "Min-Jung Jung"
date: "10/01/2021"
output:   
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This project is to create a vignette about contacting an API. I created functions to download data via interacting endpoints. I will show this process with COVID19 API. 

# Reqired pakages

I used following packages to set up function, data manipulation, and analysis with COVID API:

  * [`ggplot2`](https://ggplot2.tidyverse.org/): for creating graphics.
  * [`tidyverse`](https://www.tidyverse.org/): for data manipulation and visualization.
  * [`jsonlite`](https://cran.r-project.org/web/packages/jsonlite/): to pull data from the Covid 19 APIs.
  * [`knitr`](https://cran.r-project.org/web/packages/knitr/index.html): to generate tables.  
  * [`httr`](https://cran.r-project.org/web/packages/httr/vignettes/quickstart.html): to map closely to the underlying http protocol.
  * [`lubridate`](https://cran.r-project.org/web/packages/lubridate/index.html): for date conversion. 
  
```{r, echo=FALSE, warning = FALSE, message=FALSE, include=FALSE}
library(ggplot2)
library(tidyverse)
library(jsonlite)
library(knitr)
library(httr)
library(lubridate)
```

# Functions to contact the Covid19 Data API

To access the [Covid Data](https://documenter.getpostman.com/view/10808728/SzS8rjbc), we need to get a URL with the name of the table and attributes we want to pull from it.

Here is the base URL that I am going to use throughout defining the functions.

## Base url

```{r}
base_url = "https://api.covid19api.com"
```


I wrote several functions to contact the Covid19 API via different endpoints.

## `countryName`

This function is to generate `data.frame` of country name and Slug. In order to import different countries' datasets, I needed to type the correct country name in URL. COVID19 AIP uses Slug instead of country name. Therefore, I can use the this table to find the correct name of the country for the URL. 

```{r message=FALSE}
# Create function to find country name and slug
countryName <- function(){
  full_url = paste0(base_url,"/countries")
  country <- content(GET(url=full_url),"text")
  countrylist <- fromJSON(country)
  return(countrylist)
}
# This table will guide users to find specific country and slug.
countryName <- countryName()

```

## `covidSummary`

This function interacts with the `Summary` endpoint. It returned a `list` of 5 variables showing information such as the world total of confirmed case numbers and the world total death numbers. One of the variables was the country, which I brought out and returned a `data.frame` with the most recent data of confirmed cases, death cases, and recovered cases for each country.    

```{r message=FALSE}
covidSummary <- function(){
   full_url = paste0(base_url,"/summary")
   covid_summary_text <- content(GET(url=full_url),"text")
   covid_cases_summary_json <- fromJSON(covid_summary_text)
   # Select Country variable from json output. 
   covid_cases_summary1 <- data.frame(covid_cases_summary_json$Countries) 
   return(covid_cases_summary1)
}

``` 

## `confirmedCases`

This function interacts with the `By Country Total` endpoint. This function returns a `data.frame` of daily confirmed case numbers by a specific country during the specified dates (7/01/2020 - 09/30/2021). Users can select the country to get a different country's dataset. I chose only 4 variables to display, which I am going to mainly use. Users can use the countryName table to find countries to look up. 

```{r message=FALSE}
confirmedCases <- function(country){
    country <- tolower(country)
    country <- sub(" ", "-", country)
 
    full_url = paste0(base_url,"/total/country/",country,
                      "/status/confirmed?from=2020-07-01T00:00:00Z&to=2021-09-30T00:00:00Z")
    confirmed_cases_text = content(GET(url=full_url),"text")
    confirmed_cases_json = fromJSON(confirmed_cases_text)
    covid_confirmed_cases <- confirmed_cases_json

  if(country %in% countryName$Slug){
    covid_confirmed_cases <- covid_confirmed_cases %>% 
                           select(Country, Cases, Status, Date) # I choose 4 columns to display.
  } else { 
      message <- paste("ERROR: Argument for country was not found in the Slug column.", 
                    "Use countryName() to find the country you are looking",
                    "for and use Slug.")
    stop(message)
  }
return(covid_confirmed_cases)
}

# 1.User(s) can select different countries.
confirmed_cases <- confirmedCases("United States")
```

## `deathCases`

This function interacts with the `By Country Total` endpoint with modification of status changed to deaths. This function returns a `data.frame` of daily deaths cases number by specific country during the specified dates (7/01/2020 - 09/30/2021). Users can select country to get different country dataset. I choose only 4 variables to display, which I am going to mainly use. Users can use the countryName table to find countries you want to look up. 

```{r message=FALSE}
deathCases <- function(country){
  country <- tolower(country)
  country <- sub(" ", "-", country)
 
  full_url = paste0(base_url,"/total/country/",country,
                      "/status/deaths?from=2020-07-01T00:00:00Z&to=2021-09-30T00:00:00Z")
  deaths_cases_text = content(GET(url=full_url),"text")
  deaths_cases_json = fromJSON(deaths_cases_text)
  if(country %in% countryName$Slug){
      covid_deaths_cases <- deaths_cases_json  %>% 
                            select(Country, Cases, Status, Date) 
  } else {
    message <- paste("ERROR: Argument for country was not found in the Slug column.", 
                     "Use countryName() to find the country you are looking",
                     "for and use Slug.")
    stop(message)
  }
return(covid_deaths_cases)
}

# 2. User(s) can select different countries.
death_cases <- deathCases("United States")
```

## `recoveredCases`

This function interacts with the `By Country Total` endpoint with modification of status changed to recovered.This function returns a `data.frame` of daily recovered case numbers by specific countries during the specified dates (7/01/2020 - 09/30/2021). Users can select the country to get different countries' datasets. I chose only 4 variables to display, which I am going to primarily use. Users can use the countryName table to find countries to look up.   

```{r message=FALSE}
recoveredCases <- function(country){
  country <- tolower(country)
  country <- sub(" ", "-", country)
  # If you type in country name as slug, it will return the data correctly. 
  full_url = paste0(base_url,"/total/country/", country,
                      "/status/recovered?from=2020-07-01T00:00:00Z&to=2021-09-30T00:00:00Z")
  recovered_cases_text = content(GET(url=full_url),"text")
  recovered_cases_json = fromJSON(recovered_cases_text)
  if(country %in% countryName$Slug){
    covid_recovered_cases <- recovered_cases_json  %>% 
                           select(Country, Cases, Status, Date)
  } else {
      message <- paste("ERROR: Argument for country was not found in the Slug column.", 
                       "Use countryName() to find the country you are looking",
                       "for and use Slug.")
      stop(message)
  }
return(covid_recovered_cases)
}

# 3.User(s) can select different countries.
recovered_cases <- recoveredCases("United States")

```

```{r}
# Data manipulation

# 4.I created one dataset with two data.frame datasets by row merging.
country_all_cases <- rbind(confirmed_cases, death_cases)
```

## `confirmedCasesState`

This function interacts with the `Day One Live` endpoint with modification of status changed to confirmed. This function returns a `data.frame` of daily confirmed case numbers by specific states from 11/22/2020 to present. Users can select state names (encompassing states and other locations) to get different states' datasets. I chose only 6 variables to display, the ones I will mainly use. 

```{r message=FALSE}

confirmedCasesState <- function(state_name){
  # There are state names with one word and more than two words. You can simply type the full name of state with space between or type the abbreviated name (i.e. NC, GA, VA).  
  # I set it up to be lower case.  
  state_name <- tolower(state_name)
  
  full_url = paste0(base_url,
                    "/dayone/country/united-states/status/confirmed/live?province=",
                    state_name)
  # This function will insert %20 in the space between two words or more state name.
  URLencode(full_url)
  covid_cases_by_states_text = content(GET(url=URLencode(full_url)),"text")
  covid_cases_by_states_json = fromJSON(covid_cases_by_states_text)
  covid_cases_by_states <- covid_cases_by_states_json %>% 
                             select(Country, Province, City, Cases, Status, Date) 
  return(covid_cases_by_states)
}

# 5.User(s) can select different state names.
state_confirmedData <- confirmedCasesState("North Carolina")

# 5-1.When states have different county level information, last row is the sum of all confirmed case number of the specified state's counties.
# I am using North Carolina which has county level information. 
# Therefore, I deleted the last row to keep the state level data only.
state_confirmedData1 <- state_confirmedData %>% filter(row_number() <= n()-1)

```

## `deathsCasesState`

This function interacts with the `Day One Live` endpoint with modification of status changed to deaths. This function returns a `data.frame` of daily deaths case numbers by specific states from 11/22/2020 to present. Users can select state name to get different state datasets. I chose only 6 variables to display, the ones I will mainly use.

```{r message=FALSE}
deathsCasesState <- function(state_name){
  # There are state names with one word and more than two words. 
  # You can simply type the full name of state with space between or type the abbreviated name (i.e. NC, GA, VA).  
  # I set it up to be lower case.  
  state_name <- tolower(state_name)
  full_url = paste0(base_url,
                    "/dayone/country/united-states/status/deaths/live?province=",
                    state_name)
  URLencode(full_url)
  covid_cases_by_states_text = content(GET(url=URLencode(full_url)),"text")
  covid_cases_by_states_json = fromJSON(covid_cases_by_states_text)

  covid_cases_by_states <- covid_cases_by_states_json %>% 
                           select(Country, Province, City, Cases, Status, Date) 
  return(covid_cases_by_states)
}

# 6.User(s) can select different state names.
state_deathData <- deathsCasesState("North Carolina")

# 6-1.When states have different county level information, last row is the sum of all confirmed case number of the specified state's counties.
# I am using North Carolina which has county level information. 
# Therefore, I deleted the last row to keep the state level data only.
state_deathData1 <- state_deathData %>% filter(row_number() <= n()-1)

```

```{r}
# Data loading

# 7.I created this data set for the EDA by row merging.
state_all_cases <- rbind(state_confirmedData1, state_deathData1)

```

## `liveConfirmedCases`

This function interacts with the `Live By Country And Status After Date` endpoint of modification of date as 07/01/2021. This function returns a `data.frame` of daily confirmed cases of specified country from a set date to present. Users can select country name to get different country datasets. I chose only 6 variables to display, the ones I will mainly use.

```{r}
liveConfirmedCases <- function(country){
  # If you type in country name as slug, it will return the data correctly. 
  if(country %in% countryName$Slug){
    full_url = paste0(base_url,"/live/country/", country,
                      "/status/confirmed/date/2021-07-01T00:00:00Z")
    covid_cases_live_text = content(GET(url=full_url),"text")
    covid_cases_live_json = fromJSON(covid_cases_live_text)
    covid_state_cases <- covid_cases_live_json %>% 
                         select(Country, Province, Confirmed, Deaths, Active, Date) 
    return(covid_state_cases)
  } else {
      message <- paste("ERROR: Argument for country was not found in the Slug.", 
                       "Look up countryName to find the country you are looking",
                       "for and use Slug.")
      stop(message)
  }
}

# 8.User(s) can select different countries.
country_liveCases <- liveConfirmedCases("united-states")

```

## `dateManipulation` 

This function helps to manipulate the date column(character format) to date formation and creates one variable with time span with every quarter. 

```{r}
dateManipulation <- function(dataset){
  dataset$Date <- as.Date(dataset$Date)
  dataset <- dataset %>% 
    mutate("Time_span" = if_else(Date <= as.Date("2020-09-30"), "2020Q3",
                          if_else(Date <= as.Date("2020-12-31"), "2020Q4",
                           if_else(Date <= as.Date("2021-03-31"), "2021Q1", 
                            if_else(Date <= as.Date("2021-06-30"), "2021Q2", 
                             if_else(Date <=   as.Date("2021-09-30"),"2021Q3", "2021Q4")))))
    )
  dataset$Time_span <- as.factor(dataset$Time_span)
  return(dataset)
}              

# 9.data manipulation using `dateManipulation` function

# confirmed case data
confirmed_month <- dateManipulation(confirmed_cases)

# 10.US all cases (confirmed and death) data
us_all_cases_month <- dateManipulation(country_all_cases)

# 11.NC all cases (confirmed and death) data
state_all_cases_month <- dateManipulation(state_all_cases)

```

## `riskStatusManipulation`

This function helps to manipulate data with creating three variables. Death rate, death rate status, and risk status are the three variables. US [death rates](https://coronavirus.jhu.edu/data/mortality) is about 1.6 %. I wanted to check on how US states display death rates. Also, I created a Death Rate status variable based on the death rate and divided it into 4 categories. Lastly, I created a risk status variable based on the confirmed case numbers at the state level.  

```{r}
#I created this function for data summaries and visualization.
riskStatusManipulation <- function(dataset){
  dataset <- dataset %>% 
    mutate("DeathRate"= (Deaths/Confirmed)*100, 
           "DeathRateStatus"= if_else(DeathRate > 2, "High",
                               if_else(DeathRate > 1, "Medium", 
                                if_else(DeathRate >0.5, "Low", "Very Low"))
                               ), 
           "RiskStatus" = if_else(Confirmed > 1250000, "Very High",
                           if_else(Confirmed > 750000, "High", 
                            if_else(Confirmed > 350000, "Medium", 
                             if_else(Confirmed > 150000, "Low", "Very Low"))))
    )            
  return(dataset)
}
```

```{r}
# I filtered date to 10/01/2021 to only view the date's information.
us_newlive <- country_liveCases %>% filter(Date == "2021-10-01T00:00:00Z")

# 12.US live data using riskStatusManipulation function.
us_newlive_risk <- riskStatusManipulation(us_newlive) %>% 
                    as_tibble()
us_live_risk <- dateManipulation(us_newlive_risk)

# Overwrite RiskStatus column with factor version
us_live_risk$RiskStatus <- as.factor(us_live_risk$RiskStatus)
# Use ordered function on a factor to order the levels
us_live_risk$RiskStatus <- ordered(us_live_risk$RiskStatus, levels = c("Very Low", "Low", "Medium", "High", "Very High"))

# Overwrite DeathRateStatus column with factor version
us_live_risk$DeathRateStatus <- as.factor(us_live_risk$DeathRateStatus)
# Use ordered function on a factor to order the levels
us_live_risk$DeathRateStatus <- ordered(us_live_risk$DeathRateStatus, levels = c("Low", "Medium", "High"))

```

# Exploratory Data Analysis

I pulled the data using functions I created interacting with endpoints. After using manipulation functions and merging different datasets, I mainly selected four data sets for the exploratory analysis.

  * `country_live_risk`: Confirmed and death cases by 55 states on 10/01/2021 in US. 
  * `state_all_cases_month`: Confirmed cases in 101 NC Counties from 11/22/2020 to present.
  * `confirmed_month`: US confirmed cases from 07/01/2020 to 09/30/2021.
  * `us_all_cases_month`: US confirmed and death cases merged from 07/01/2020 to 09/30/2021.  

Let's talk about the Covid-19 pandemic in the U.S and how confirmed cases and death cases change over time. In order to see the change patterns of both confirmed and death cases, I created a scatter plot by date. 


```{r message=FALSE}

us_all_cases_month_wide <- us_all_cases_month %>%
                            pivot_wider(names_from = "Status",
                                        values_from = "Cases")

# Scatter plot confirmed cases in US
ggplot(data = us_all_cases_month_wide, aes(x = Date, 
                                      y = confirmed)) +
  geom_point(alpha=0.50) + 
   
  theme(axis.text.x = element_text(angle = 45,hjust=1)) + 
  ggtitle("US confirmed cases vs deaths cases") + 
  geom_smooth(method = lm, color = "blue")  

# Scatter plot deaths cases in US
ggplot(data = us_all_cases_month_wide, aes(x = Date, 
                                           y = deaths)) +
  geom_point(alpha=0.50) + 
  theme(axis.text.x = element_text(angle = 45,hjust=1)) +
  ggtitle("US deaths cases only") + 
  geom_smooth(method = lm, color = "blue")  

```
In the confirmed scatter plot, we see that the cases increased rapidly at the end of 2020 and beginning of 2021, and then it slowed down. However, it started increasing again since August 2021. It may have been related to mask policies, vaccination, or seasonal changes, but the fact is that covid cases are increasing. Similar to confirmed cases, deaths cases also show similar pattern. Deaths cases are increasing over time, but it slowed  down around March of 2021. It looks like deaths cases are increasing again around September. It will be interesting to see if we have a similar pattern of winter of 2020. 


Now, I wanted to see how the confirmed cases are increasing by every quarter using side by side box plots. 

```{r message=FALSE}
# Box plot
us_confirmed_states<- us_all_cases_month %>% 
                        filter(Status == "confirmed") 
#Box plot for confirmed cases by time line.
ggplot(us_confirmed_states, aes(x = Time_span, 
                                y= Cases, 
                                fill=Time_span)) +
  geom_boxplot() +
  scale_x_discrete("Time Span") +
  ggtitle("US Covid confirmed cases comparison by time span") +
  scale_fill_brewer(palette="Dark2") +
  scale_fill_discrete(name = "Time Span", 
                      labels = c("2020Q3" = "2020 Q3", 
                                 "2020Q4" = "2020 Q4", 
                                 "2021Q1" = "2021 Q1", 
                                 "2021Q2" = "2021 Q2", 
                                 "2021Q3" = "2021 Q3")
                                ) 
```

Similar to the scatter plot of confirmed cases, the median line of 2021 quarter 1 is a lot higher than 2020 quarter 4. Also, the interquartile range of 2020 quarter 4 is big, which means that during 2020 quarter 4, case numbers rapidly increased. 

Now, let's look at how death cases are increasing by quarter. 

```{r message=FALSE}
#Setting the data set with only deaths status
us_deaths_states<- us_all_cases_month %>% 
                    filter(Status == "deaths") 
#Box plot for death cases by timeline
ggplot(us_deaths_states, aes(x = Time_span, 
                                y= Cases, 
                                fill=Time_span)
                            ) +
  geom_boxplot() +
  scale_x_discrete("Time Span") +
  ggtitle("US Covid deaths cases comparison by time span") +
  scale_fill_brewer(palette="BuPu") +
  scale_fill_discrete(name = "Time Span", 
                      labels = c("2020Q3" = "2020 Q3", 
                                 "2020Q4" = "2020 Q4", 
                                 "2021Q1" = "2021 Q1", 
                                 "2021Q2" = "2021 Q2", 
                                 "2021Q3" = "2021 Q3")
                                ) 

```

I also created side by side box plots for the death cases, and death cases increased rapidly between the last quarter of 2020 and the first quarter of 2021. 
 
Now, let's move down to the state level, beginning with discussing some summary statistics. 

```{r}
# Summary table for US states
us_bystate <- us_live_risk %>% 
                summarise(Min = min(Confirmed), 
                          Median = median(Confirmed),
                          Average = mean(Confirmed),
                          Max = max(Confirmed),
                          IQR = IQR(Confirmed)
                          )

# Display a table of the summary stats.
kable(us_bystate, caption = "Summary Stats by US states")
```

First, `us_bystate` summary table using `us_live_risk` data shows interesting information. The average confirmed cases (790,137.1) is a little bit higher than the median score (508494), and the IQR score is also not extremely high. However, the difference between minimum (269) and maximum (4,718,816) numbers is extremely big. If I didn't check the Minimum score or Maximum score, I would miss the gap between the two end values. 

```{r}
# Summary stats for two categorical variables
Summary_us_risk <- us_live_risk %>% 
                    group_by(RiskStatus) %>% 
                    summarise(Average = mean(Confirmed),
                              Median = median(Confirmed), 
                              IQR = IQR(Confirmed)
                              )

# Display a table of the summary stats.
kable(Summary_us_risk, caption = "Summary stats by risk status")

```

`Summary_us_risk` table shows the average, median, and IQR scores of each risk state. I created a variable called RiskStatus based on the mean and median score. Five status set points of the number of confirmed cases are from "Very Low" (under 150,000 cases) to "Very High"(higher than 1,250,000 cases). Similar to the first summary table,  this table also shows interesting results on very high status scores. The average score is significantly higher than the median score, and the IQR score in the very high category is also large. This means that states with confirmed cases higher than 1,250,000 are more spread out, but the majority are still around the median scores.  

```{r}
# Arrange data by descender order
us_live_risk1 <- us_live_risk %>% 
                  arrange(desc(Confirmed))

# Select 10 top rows
top10states <- us_live_risk1[1:10,]

#Top10 states by risk status
state_highrisk <- table(top10states$Province, top10states$RiskStatus)
kable(state_highrisk, 
      caption = "Top 10 States of risk status")
```

`state_highrisk` contingency table shows where the top 10 states with the most confirmed cases belong in terms of their risk status. Nine out of 10 states are in the very high risk group, and North Carolina is one of the states in the high risk group. One thing we should remember is that this grouping is not based on the population. It directly looks at the confirmed cases only. Therefore, if considering state populations, risk categories may change.    

```{r}
#Bar plot for top 10 states
ggplot(data=top10states, aes(x=Province, 
                             y=Confirmed)) +
  geom_bar(stat="identity", fill="orange") +
  labs(x = "Top 10 States", 
       title = "Top 10 US states on confirmed cases") +
  geom_text(aes(label=Confirmed), 
            vjust=1.6, 
            color="black", 
            size=3.5) +
  theme(axis.text.x = element_text(angle = 45, hjust=1))

```

While state_highrisk contingency table shows where each states belongs in terms of risk category, the bar plot shows the number of confirmed cases. We now know that California is the top state with the most confirmed cases followed by Texas, Florida, and New York.   

Now, let's look at both confirmed cases and death cases and their relationship. Are deaths higher in the states with higher confirmed cases? We can use the correlation score to see the relationship. 

```{r message=FALSE}
# Correlation summary
corr_us <- cor(us_live_risk$Confirmed, us_live_risk$Deaths)

# Display correlation summary
kable(corr_us, col.names = c("Correlation"))

#Scatter plot: confirmed cases by death cases
ggplot(data = us_live_risk, aes(x = Confirmed, 
                                y = Deaths)) +
  geom_point() + 
  theme(axis.text.x = element_text(angle = 45,hjust=1)) + 
  ggtitle("Correlation between deaths and confirmed cases") + 
  geom_smooth(method = lm, color = "blue")  
```

The correlation between confirmed cases and death cases is 0.975, which is really high being close to 1. The scatter plot shows the pattern of the relationship, as well. Based on the results, the top 10 confirmed case states are more likely to show higher death cases. 

The bar plot below shows the top 10 states on death cases. 

```{r}
# Sub setting top 10 states of deaths cases.
# Descender order
us_live_death1 <- us_live_risk %>% 
                    arrange(desc(Deaths))

# Filter top 10 states
top10statesdeath <- us_live_death1[1:10,]

# Bar plot for top 10 states by deaths cases.
ggplot(data=top10statesdeath, aes(x=Province, 
                                  y=Deaths)) +
  geom_bar(stat="identity", fill="orange") +
  labs(x = "State", title = "Top 10 US states on deaths cases") +
  geom_text(aes(label=Deaths), vjust=1.6, color="black", size=3.5) +
  theme(axis.text.x = element_text(angle = 45, hjust=1))

```

As the prediction of the correlation between confirmed cases and death cases, California is again the top state showing the highest number of deaths followed by Texas, Florida, and New York. Interestingly, Death cases in New York are similar to Florida, which infers that the death rate in New York is higher. Let's move on to see the death rate of the top 10 states. 

```{r}
# Sub setting top 10 states by deaths rate.
# Descender order
us_live_deathrate <- us_live_risk %>% 
                      arrange(desc(DeathRate))

# Filter top 10 states and select Province and DeathRate columns. 
top10statesDeathRate <- us_live_deathrate[1:10,] %>% 
                          select (Province, DeathRate)

# Display top 10 states by death rate.
kable(top10statesDeathRate, 
      col.names = c("States", "Death Rate"),
      caption = "Death rate by top 10 states", digits=2)

```

`top10statesDeathRate` table shows top 10 states for death rates. Surprisingly, California and Texas are not on this top 10 list. New Jersey, Massachusetts, and New York states ranked in the top 3. These results show more dynamics between confirmed cases and deaths cases, and the death rates do not always go by the number of cases. Then, it will be interesting to see the relationship between deathrate and confirmed cases. I created `DeathRateStatus` variable to see how many states are in high, medium, and low death rate groups. Therefore, instead of directly comparing the two numbers, I used categorical variables (RiskStatus and DeathRateStatus) to see how many states belong to each group in the contingency table.   

```{r}
# Contingency table 

# United Status
usDeathRate_status <- table(us_live_risk$RiskStatus, us_live_risk$DeathRateStatus)

# change row names.
rownames(usDeathRate_status) <- c("VeryLow risk", "Low risk", 
                                  "Medium risk", "High risk", 
                                  "VeryHigh risk")

# Display contingency table.
kable(usDeathRate_status, 
      col.names = c('Low deathrate', 'Medium deathrate', 'High deathrate'),
      caption = "US states: Risk status vs death rate status")
```

When looking at the contingency table with confirmed cases risk groups and death rate groups, it still shows consistent results. The very high risk category in terms of confirmed cases grouped in the medium and high death rate categories. Similarly, the very low and low risk category in terms of the confirmed cases grouped in the low and medium death rate categories.       

This bar plot shows the contingency table more visually. 

```{r}
#Bar plot of risk status in 55 states

ggplot(data = us_live_risk, aes(x=RiskStatus)) +
  geom_bar(aes(fill = as.factor(DeathRateStatus))) + 
  labs(x = "Risk Status", 
       title = "Risk status in 55 states") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  scale_fill_discrete(name = "Death Rate Status") 
              

```

You will see that the low death rate states are mainly in medium to low and very low risk groups (low confirmed cases) and high death rate states are in the medium to high and very high risk groups (high confirmed cases). 

Lastly, I wanted to see how confirmed cases and death cases are changing in wake county.  

```{r message=FALSE}
# sub setting rows of Wake county
Wake_cases <- state_all_cases_month %>% 
                filter(City =="Wake") 

# Making status column to two columns (deaths and confirmed)
wake_cases_wide <- Wake_cases %>% 
                    pivot_wider(names_from = "Status", values_from = "Cases")

# Scatter plots of confirmed cases for wake county
ggplot(data = wake_cases_wide, aes(x = Date, y = confirmed)) +
  geom_point(alpha=0.50) + 
  theme(axis.text.x = element_text(angle = 45,hjust=1)) + 
  ggtitle("Wake county confirmed cases by date") +
  geom_smooth(method = lm, color = "blue") +
  labs(x = "Month in 2021")

# Scatter plots of death cases for wake county
ggplot(data = wake_cases_wide, aes(x = Date, y = deaths)) +
  geom_point(alpha=0.50) + 
  theme(axis.text.x = element_text(angle = 45, hjust=1)) + 
  ggtitle("Wake county deaths cases") +
  geom_smooth(method = lm, color = "blue") +
  labs(x = "Month in 2021", y = "Deaths cases")

```

Looking at scatter plots, the confirmed scatter plot reveals that in the beginning of 2021 the cases increased rapidly. Then the cases increased rapidly again since August. One possible explanation is due to wake county school stared. Looking at the deaths cases scatter plot, the deaths cases also increased as confirmed cases increased. Especially, there was a big jump on the deaths cases in March from under 600 cases to almost 700 cases. However, Since April it has been slow down, which is a good news. Since flu season is coming, we will need to pay attention how the pattern will change. 

These histograms can also show how quickly confirmed cases and death cases are changing by frequency of the same case numbers. 

```{r message=FALSE}
# Histogram plot

# Wake county confirmed cases in 2021
wake_confirmed_cases <- state_all_cases_month %>% 
                          filter(Status == "confirmed") %>% 
                          filter(City == "Wake")

# Wake county deaths cases in 2021
wake_deaths_cases <- state_all_cases_month %>% 
                        filter(Status == "deaths") %>% 
                        filter(City == "Wake")

# Histogram of confirmed cases in 2021
ggplot(wake_confirmed_cases, aes(Cases)) + 
  geom_histogram(color = "brown4", fill = "brown1") +
  labs(x = "Confirmed Cases", 
       title = "Confrimed cases in Wake county in 2021")

# Histogram of deaths cases in 2021
ggplot(wake_deaths_cases, aes(Cases)) + 
  geom_histogram(color = "darkorange4", fill = "darkorange1") +
  labs(x = "Deaths Cases", 
       title = "Deaths cases in Wake county in 2021")

```

The confirmed histogram shows how many confirmed cases were in Wake County by number of counts (days), so the longer the number of cases remained the same, the larger the bar on the histogram, which means the slower the increase in number of cases. In other words, in the confirmed histogram, the highest bar is at 80,000 confirmed cases. The time period when the confirmed cases were 80,000 was most frequent, and we can interpret this as during those 62 counts (62 days), the confirmed cases increased slowly, which occurred around the second quarter of the year. However, confirmed cases are rapidly increasing again because the histogram shows that later this year, the frequency of confirmed cases at a given number count was much lower, meaning that the number of counts with that number of cases was smaller. This shows that the cases changed more quickly. 

In the deaths histogram, death cases stagnated between 730 and 750, which means that there were a higher number of days in which the total death cases remained about the same. In other words, death cases are more slowly increasing in the later months of 2021.  It is good to know that in the midst of the more quickly increasing of confirmed cases, death cases are slowing down in Wake county.  

# Wrap-Up

In this vignette, I built several functions to interact with some of the COVID19 API’s endpoints to retrieve data for the exploratory analysis. I used tables, numerical summaries, and four different types of graphs to explain what I found. With an analysis, I found that Covid confirmed cases are still increasing with death cases, but there are variations between states in the U.S. Sadly, North Carolina is one of the top 10 states of confirmed cases. Also, confirmed cases are increasing fast in wake county. 

I selectively focused on the US cases in this vignette because US is the top country of all the confirmed cases, but it will be very interesting to see how other countries are doing navigating functions I created here. 

