---
title: "covid-vignette-api"
author: "Min-Jung Jung"
date: "9/20/2021"
output:   
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This project is to create a vignette about contacting an API. I created functions to download data via interacting endpoints. I will show this process with COVID API. 


# Reqired pakages

I used following packages to set up function, data manipulation, and analysis with COVID API:

  * `knitr`: to generate pretty tables of date using the `kable()` function.  
  * `tidyverse`: to manipulate data, generate plots (via `ggplot2`), and to use piping/chaining.
  * `rmarkdown`: to knit output files manually using the `render()` function.  
  * `jsonlite`: to pull data from various endpoints of the Covid 19 APIs.
  
```{r, echo=TRUE, warning = FALSE, message=FALSE, include=FALSE}
library(ggplot2)
library(tidyverse)
library(jsonlite)
library(knitr)
library(httr)
library(lubridate)
```

# JSON Data

JSON, also known as [*Java Script Object Notation*](https://www.json.org/json-en.html) is a text based format for storing and/or transporting data. JSON is used across the internet and databases because of its text based format for storing data. In addition, JSON can represent 2D data, [hierarchical data](https://en.wikipedia.org/wiki/Hierarchical_database_model), and use key-value pairs.

Since JSON is a text based format, we need to load in R packages that can handle the text based format.

# Packages used for reading JSON data in R

Here are R packages that can be used to work with JSON data:

* `rjson`
* `RJSONIO`
* `jsonlite`

For this project, `rjson` and `jsonlite` is used.

Couple resources for understanding the functionality of `rjson` and `jsonlite`:

1. [https://cran.r-project.org/web/packages/rjson/rjson.pdf](https://cran.r-project.org/web/packages/rjson/rjson.pdf)

2. [https://cran.r-project.org/web/packages/jsonlite/jsonlite.pdf](https://cran.r-project.org/web/packages/jsonlite/jsonlite.pdf)

# Contact the Covid Data API

To access the [Covid Data](https://documenter.getpostman.com/view/10808728/SzS8rjbc), we need to get a URL with the name of the table and attributes we want to pull from it.

In addition, I wrote 6 more functions that take different endpoints.

# Base url

```{r}
base_url = "https://api.covid19api.com"
```

## country name

#This helper function is to generate data.frame of country name and Slug. 

```{r}
country_Name <- function(){
  country <- GET("https://api.covid19api.com/countries")
  countrylist <- fromJSON(rawToChar(country$content))
  countrylist1 <- as_tibble(data.frame(Country = countrylist$Country, Slug = countrylist$Slug))
  return(countrylist1)
}

countryName <- country_Name()
```


## Summary data

#This function interacts with the `Summary` endpoint. 

```{r}
covid_summary_cases <- function(){
   full_url = paste0(base_url,"/summary")
   covid_summary_text <- content(GET(url=full_url),"text")
   covid_cases_summary_json <- fromJSON(covid_summary_text)
   covid_cases_summary1 <- data.frame(covid_cases_summary_json$Countries) 
   return(covid_cases_summary1)
}

covidSum <- covid_summary_cases()

``` 

## confirmed cases
#This function interacts with the `By Country Total` endpoint. 

```{r}
get_confirmed_cases <- function(country){
     if(country %in% countryName$Slug){
  full_url = paste0(base_url,"/total/country/",country,"/status/confirmed?from=2020-07-01T00:00:00Z&to=2021-09-30T00:00:00Z")
  confirmed_cases_text = content(GET(url=full_url),"text")
  confirmed_cases_json = fromJSON(confirmed_cases_text)
      covid_confirmed_cases <- confirmed_cases_json %>% select(Country, Cases, Status, Date) #%>% rename(ConfirmedCases =  Cases) 
  return(covid_confirmed_cases)
     }
        else {
      message <- paste("ERROR: Argument for country was not found in the Slug.", 
                       "Look up countryName to find the country you are looking for and use Slug.")
      stop(message)
    }
}

confirmed_cases <- get_confirmed_cases("united-states")

```

# Death cases

#This function interacts with the `By Country Total` endpoint with status changed to deaths. 

```{r}
get_deaths_cases <- function(country){
    if(country %in% countryName$Slug){
  full_url = paste0(base_url,"/total/country/",country,"/status/deaths?from=2020-07-01T00:00:00Z&to=2021-09-30T00:00:00Z")
  deaths_cases_text = content(GET(url=full_url),"text")
  deaths_cases_json = fromJSON(deaths_cases_text)
       covid_deaths_cases <- deaths_cases_json  %>% select(Country, Cases, Status, Date) #%>% rename(DeathsCases =  Cases) 
  return(covid_deaths_cases)
    }
        else {
      message <- paste("ERROR: Argument for country was not found in the Slug.", 
                       "Look up countryName to find the country you are looking for and use Slug.")
      stop(message)
    }
}
death_cases <- get_deaths_cases("united-states")

```

# Recovered cases

#This function interacts with the `By Country Total` endpoint with status changed to recovered.  

```{r}
get_recovered_cases <- function(country){
  if(country %in% countryName$Slug){
  full_url = paste0(base_url,"/total/country/", country,"/status/recovered?from=2020-07-01T00:00:00Z&to=2021-09-30T00:00:00Z")
  recovered_cases_text = content(GET(url=full_url),"text")
  recovered_cases_json = fromJSON(recovered_cases_text)
       covid_recovered_cases <- recovered_cases_json  %>% select(Country, Cases, Status, Date) #%>% rename(RecoveredCases =  Cases)
  return(covid_recovered_cases)
  }
      else {
      message <- paste("ERROR: Argument for country was not found in the Slug.", 
                       "Look up countryName to find the country you are looking for and use Slug.")
      stop(message)
    }
}
recovered_cases <- get_recovered_cases("united-states")

```

```{r}
us_all_cases <- rbind(confirmed_cases, death_cases)

```

## Us confirmed cases by state

#This function interacts with the `Day One Live` endpoint with status changed to confirmed

```{r}
confirmed_cases_bystate <- function(state_name){
  state_name <- tolower(state_name)
  two_word_states = list("new hampshire", "new jersey", "new mexico","new york","north carolina","north dakota","south carolina","south dakota", "distrct of columbia", "puerto rico","Northern Mariana Islands", "Virgin Islands", "Rhode Island")
  if (state_name %in% two_word_states){
     full_url = paste0(base_url,"/dayone/country/united-states/status/confirmed/live?province=",state_name)
     URLencode(full_url)
    covid_cases_by_states_text = content(GET(url=URLencode(full_url)),"text")
    covid_cases_by_states_json = fromJSON(covid_cases_by_states_text)
  }
    else{
    full_url = paste0(base_url,"/dayone/country/united-states/status/confirmed/live?province=",state_name)
    covid_cases_by_states_text = content(GET(url=full_url),"text")
    covid_cases_by_states_json = fromJSON(covid_cases_by_states_text)
    }
     covid_cases_by_states <- covid_cases_by_states_json %>% select(Country, Province, City, Cases, Status, Date) 
  return(covid_cases_by_states)
}

nc_confirmedData <- confirmed_cases_bystate("North Carolina")
nc_confirmedData1 <- nc_confirmedData %>% filter(row_number() <= n()-1)
```

## Us Death cases by state

#This function interacts with the `Day One Live` endpoint with status changed to deaths.

```{r}
deaths_cases_bystate <- function(state_name){
  state_name <- tolower(state_name)
  two_word_states = list("new hampshire", "new jersey", "new mexico","new york","north carolina","north dakota","south carolina","south dakota", "distrct of columbia", "puerto rico","Northern Mariana Islands", "Virgin Islands", "Rhode Island")
  if (state_name %in% two_word_states){
     full_url = paste0(base_url,"/dayone/country/united-states/status/deaths/live?province=",state_name)
     URLencode(full_url)
    covid_cases_by_states_text = content(GET(url=URLencode(full_url)),"text")
    covid_cases_by_states_json = fromJSON(covid_cases_by_states_text)
  }
    else{
    full_url = paste0(base_url,"/dayone/country/united-states/status/deaths/live?province=",state_name)
    covid_cases_by_states_text = content(GET(url=full_url),"text")
    covid_cases_by_states_json = fromJSON(covid_cases_by_states_text)
    }
     covid_cases_by_states <- covid_cases_by_states_json %>% select(Country, Province, City, Cases, Status, Date) 
  return(covid_cases_by_states)
}

nc_deathData <- deaths_cases_bystate("North Carolina")
nc_deathData1 <- nc_deathData %>% filter(row_number() <= n()-1)

```
```{r}
#created one data set merging confirmed and death data sets
nc_all_cases <- rbind(nc_confirmedData1, nc_deathData1)

```

## live cases by country

endpoint `Live By Country And Status After Date`

```{r}
live_cases <- function(country){
       if(country %in% countryName$Slug){
   full_url = paste0(base_url,"/live/country/",country,"/status/confirmed/date/2021-07-01T00:00:00Z")
    covid_cases_live_text = content(GET(url=full_url),"text")
    covid_cases_live_json = fromJSON(covid_cases_live_text)
      covid_state_cases <- covid_cases_live_json  %>% select(Country, Province, Confirmed, Deaths, Active, Date) #%>% rename(RecoveredCases =  Cases)
      return(covid_state_cases)
       }
      else {
      message <- paste("ERROR: Argument for country was not found in the Slug.", 
                       "Look up countryName to find the country you are looking for and use Slug.")
      stop(message)
    }
}

us_liveCases <- live_cases("united-states")

```

## Date manipulation

#This function helps to manipulate date column to date formation and create one variable with time span by quater. 

```{r}
get_date_case <- function(dataset){
dataset$Date <- as.Date(dataset$Date)
dataset <- dataset %>% 
                  mutate("Time_span" = if_else(Date <= as.Date("2020-09-30"), "2020Q3",
                                        if_else(Date <= as.Date("2020-12-31"), "2020Q4",
                                          if_else(Date <= as.Date("2021-03-31"), "2021Q1", 
                                            if_else(Date <= as.Date("2021-06-30"), "2021Q2", 
                                              if_else(Date <= as.Date("2021-09-30"),"2021Q3","2021Q4")))))
                         )

                  dataset$Time_span <- as.factor(dataset$Time_span)
  return(dataset)
}              

confirmed_month <- get_date_case(confirmed_cases)

us_all_cases_month <- get_date_case(us_all_cases)

nc_all_cases_month<- get_date_case(nc_all_cases)

```


```{r}
#As of July 28,2020 CDC covid case map
#https://twitter.com/cdcgov/status/1288609081696169990

#we can create a function that will manipulate our data to prepare for data summaries and visualization
usRiskyData <- function(dataset){
  dataset <- dataset %>% 
            mutate("DeathRate"= (Deaths/Confirmed)*100, 
                   "DeathRateStatus"= if_else(DeathRate > 2, "4.High",
                                  if_else(DeathRate > 1, "3.Medium", 
                                          if_else(DeathRate >0.5, "2.Low", "1.Very low"))), 
                   "RiskStatus" = if_else(Confirmed > 1250000, "5.Veryhigh",
                                    if_else(Confirmed > 750000, "4.High", 
                                      if_else(Confirmed > 350000, "3.Medium", 
                                        if_else(Confirmed > 150000, "2.Low", "1.Very Low"))))
                                  )            
  return(dataset)
}
```
```{r}
#need to change the name of file later.
us_newlive <- us_liveCases %>% filter(Date == "2021-10-01T00:00:00Z")

us_newlive_risk <- usRiskyData(us_newlive) %>% as_tibble()
us_live_risk <- get_date_case(us_newlive_risk)
```

# Exploratory Data Analysis

## Contigency Tables
```{r}
# Contingency table 
#United Status
usDeathRate_status <- table(us_live_risk$RiskStatus, us_live_risk$DeathRateStatus)
usDeathRate_status
```

```{r}

us_live_risk1 <- us_live_risk %>% arrange(desc(Confirmed))
us_live_risk2 <- us_live_risk1[1:10,]
#Top10 states by death rate status
state_deathrate <- table(us_live_risk2$Province, us_live_risk2$DeathRateStatus)
state_deathrate
```

## Numerical Summaries
```{r}
#Summary table for us states
us_bystate <- us_live_risk %>% summarise(Min = min(Confirmed), Max = max(Confirmed), Avg = mean(Confirmed), Med = median(Confirmed), IQR = IQR(Confirmed))
us_bystate
#summary by status(deaths vs confirmed)
Summary_us_all <- us_all_cases_month %>% group_by(Status) %>% summarise(Avg = mean(Cases), Med = median(Cases), IQR = IQR(Cases)) 
Summary_us_all
#two categorical variables
Summary_us_risk <- us_live_risk %>% group_by(RiskStatus) %>% summarise(Avg = mean(Confirmed), Med = median(Confirmed), IQR = IQR(Confirmed), Var = var(Confirmed)) 
Summary_us_risk
#two categorical variables
Summary_us_DR <- us_live_risk %>% group_by(DeathRateStatus) %>% summarise(Avg = mean(Confirmed), Med = median(Confirmed), IQR = IQR(Confirmed), Var = var(Confirmed))
Summary_us_DR
#correlation
corr_us <- cor(us_live_risk$Confirmed, us_live_risk$Deaths)
corr_us
```

## Graphical Summaries

```{r message=FALSE}
#Scatter plot US COVID-19 confirmed cases by time
ggplot(data = confirmed_month, aes(x = Date, y = Cases))+
  geom_point() + theme(axis.text.x = element_text(angle = 45,hjust=1)) + 
  ggtitle("Scatterplot of Confirmed Covid Cases by Confirmed Covid Cases") + geom_smooth(method = lm, color = "blue")  
```

```{r message=FALSE}
#Scatter plot: confirmed cases by death cases
ggplot(data = us_live_risk, aes(x = Confirmed, y = Deaths))+
  geom_point() + theme(axis.text.x = element_text(angle = 45,hjust=1)) + 
  ggtitle("Scatterplot of correlation between death Cases and confirmed cases") + geom_smooth(method = lm, color = "blue")  
```

```{r}
#Subsetting data by specific date and confirmed status
top_states<- nc_all_cases_month %>% filter(Status == "confirmed") %>% filter(Date == "2021-10-01") %>% arrange(desc(Cases))
top_10counties <- top_states[1:10,] #Only using top 10 confirmed counties
#Bar plot for top 10 counties
ggplot(data=top_10counties, aes(x=City, y=Cases)) +
  geom_bar(stat="identity", fill="orange") +
  labs(x = "City", title = "Top 10 Counties on Confirmed case") + geom_text(aes(label=Cases), vjust=1.6, color="black", size=3.5) +
  theme(axis.text.x = element_text(angle = 45, hjust=1))
```

```{r}
#Bar plot of risk status in 55 states

ggplot(data = us_live_risk, aes(x=RiskStatus)) +
  geom_bar(aes(fill = as.factor(DeathRateStatus))) + labs(x = " Risk Status", title = "Bar plot of Risk status in 55 states") + theme(axis.text.x = element_text(angle = 45, hjust=1)) + scale_fill_discrete(name = "DeathRate Status")
```


```{r message=FALSE}
# Scatter plot for US
ggplot(data = us_all_cases_month, aes(x = Date, y = Cases))+
  geom_point() + facet_wrap(~ Status) + theme(axis.text.x = element_text(angle = 45,hjust=1)) + 
  ggtitle("Scatterplot of U.S. Cases by date") + geom_smooth(method = lm, color = "blue")  
```

```{r message=FALSE}
# subsetting rows by two counties
Wake_cases <- nc_all_cases_month %>% filter(City=="Wake") 
Mecklenburg_cases <- nc_all_cases_month %>% filter(City=="Mecklenburg") 

# Scatter plots for wake and Meck country
ggplot(data = Wake_cases, aes(x = Date, y = Cases))+
  geom_point() + facet_wrap(~ Status) + theme(axis.text.x = element_text(angle = 45,hjust=1)) + 
  ggtitle("Scatterplot of  Wake county Cases by date") + geom_smooth(method = lm, color = "blue")  

ggplot(data = Mecklenburg_cases, aes(x = Date, y = Cases))+
  geom_point() + facet_wrap(~ Status) + theme(axis.text.x = element_text(angle = 45,hjust=1)) + 
  ggtitle("Scatterplot of  Mecklenburg county Cases by date") + geom_smooth(method = lm, color = "blue") 

```

```{r}
# Box plot
us_confirmed_states<- us_all_cases_month %>% filter(Status == "confirmed") 
#Box plot for confirmed cases by time line.
boxplot(Cases~Time_span,data=us_confirmed_states, main="Us Covid confirmed Cases Comparison by Time span using Box plot",xlab="Timespan", ylab="Cases",col=(c("blue","orange","green","gold","purple")))
```
```{r}
#Setting the data set with only deaths status
us_deaths_states<- us_all_cases_month %>% filter(Status == "deaths") #Box plot for death cases by timeline
boxplot(Cases~Time_span,data=us_deaths_states, main="Us Covid death Cases Comparison by Time span using Box plot",xlab="Timespan", ylab="Cases",col=(c("blue","orange","green","gold","purple")))
```


```{r, message=FALSE}
# Histogram plot

#need to change title and name of the x axis

ncst_confirmed_cases <- nc_all_cases_month %>% filter(Status == "confirmed") %>% filter(City == "Wake")
ncst_deaths_cases <- nc_all_cases_month %>% filter(Status == "deaths") %>% filter(City == "Wake")

ggplot(ncst_confirmed_cases, aes(Cases)) + geom_histogram(color = "blue", fill = "lightblue")
ggplot(ncst_deaths_cases, aes(Cases)) + geom_histogram(color = "blue", fill = "lightblue")

```

